Complete Liver Model for Real-time Simulation

### 2013-02-28 Doc/Miccai2013/miccai2013.tex
few comments
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@522 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
%%%

@@

\begin{abstract}
This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver.
We show that the approach is able to reproduce the deformations observed at the local level.
Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

@@

Medical simulation is expected to play an integral part in many aspects of
medical practice. From diagnosis, through planning or training to computer
assisted intervention. A good medical model has to be simple yet efficient
to be really usable.
The liver is usually modeled as a homogeneous isotropic (visco-)elastic
material \TG{refs}.% It is however non-homogeneous and modeling it as
%homogeneous necessarily requires complex non-linear material descriptions.
Recently
Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
structures on the liver deformation.

@@

It is known that the parenchyma exhibits viscoelastic behaviour \TG{Cite
something}, we however employ simpler elastic model as we are not so much
interested in time-dependent behaviour but rather in static equilibrium
under certain conditions.
%
The parenchyma is modeled using corotational finite elements. The corotational
formulation allows us to deal with geometrical non-linearities but still
rely on linear stress-strain relationship. Thus rotations and large
deformations are possible. In the corotational formulation the deformation
of each element is expressed in the local frame of reference. There are
several ways of computing the corotational frame for elements, we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.
The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. Where the
components of $\Mat{K}$ depend on the orientation of each element in the
current simulation step.

@@

\TG{add a ref. or dive into the definition? \cite{Felippa2003}}

@@

Without the loss of generality we can assume the tetrahedron consists of
nodes $p_1, p_2, p_3$ and $p_4$ and the boundary triangle has nodes $p_1, p_2$
and $p_3$. We can reorder the degrees of freedom so that the stiffness
matrix $\Mat{K}^t$ for the tetrahedron can be written as:

@@

We also combined our composite model with vascularization method proposed
by Peterl\'{i}k et al. \cite{Peterlik2012}. The method is based on linear beams
with local frames of reference (in a many aspects similar to the
corotational formulation). As such the model also handles geometric
non-linearities in the deformation. The proposed model assumes no relative
motion of vessels to the parenchyma. Segmented vascular network is
discretized into the set of linear elements that are then projected into
the tetrahedral mesh of the parenchyma. During the simulation forces and
torques resulting from the deformations of vessels by parenchyma are
computed and propagated back onto the tetrahedral mesh.

@@

@@

@@

\subsection{Global Deformations}

>>>>

%%% Comments
\usepackage{color}
\newcommand{\TG}[1]{{\color{blue}\textbf{TG: #1}}}
\newcommand{\CD}[1]{{\color{green}\textbf{CD: #1}}}
%%%

@@

\begin{abstract}
This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver (\CD{based on a coupling ?})
We show that the approach is able to reproduce the deformations observed at the local level.
Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

@@

Medical simulation is expected to play an integral part in many aspects of
medical practice. From diagnosis, through planning or training to computer
assisted intervention. A good medical model has to be simple yet efficient
to be really usable.
\CD{develop the applications for which it has to be real-time...}

\CD{First... explain the biomechanics of the liver (for instance: 
\textit{
The liver is a complex organ composed by three main structures: the parenchyma (soft tissue, that have a hyper-visco-elastic behavior) is coupled with a dense vascular network (tree-like tubular structure with a stiffer behavior) and is surrounded by the Glisson capsule  (a stiff collagenous membrane).
It illustrates the need of relying on both volume, surface and line finite elements and the need of proposing coupling methods between them.}
 }

\CD{Yet, many work on liver modeling only consider the parenchyma...}
The liver is usually modeled as a homogeneous isotropic (visco-)elastic
material \TG{refs}.% It is however non-homogeneous and modeling it as
%homogeneous necessarily requires complex non-linear material descriptions.
Recently
Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
structures on the liver deformation.

@@

\CD{This paragraph may change if we use an other model... but maybe we should directly say that we rely on vascularized model ?}
It is known that the parenchyma exhibits viscoelastic behaviour \TG{Cite
something}, we however employ simpler elastic model as we are not so much
interested in time-dependent behaviour but rather in static equilibrium
under certain conditions.
%
The parenchyma is modeled using corotational finite elements. The corotational
formulation allows us to deal with geometrical non-linearities but still
rely on linear stress-strain relationship. Thus rotations and large
deformations are possible. In the corotational formulation the deformation
of each element is expressed in the local frame of reference. There are
several ways of computing the corotational frame for elements, we rely on
the geometrical method proposed in \cite{Nesme2005}.
% NOTE: This description of corotational method is very simplified and could be extended.
The stiffness matrix $\Mat{K}$ can be expressed in terms of the deformation
$u$ and the force as $\Vec{f} = \Mat{K}(\Vec{u}) \Vec{u}$. Where the
components of $\Mat{K}$ depend on the orientation of each element in the
current simulation step.

@@

\TG{add a ref. or dive into the definition? \cite{Felippa2003}} \CD{both...}

@@

\CD{The following may appear as "trivial"... maybe we can remove this part at the end if we need space} 
Without the loss of generality we can assume the tetrahedron consists of
nodes $p_1, p_2, p_3$ and $p_4$ and the boundary triangle has nodes $p_1, p_2$
and $p_3$. We can reorder the degrees of freedom so that the stiffness
matrix $\Mat{K}^t$ for the tetrahedron can be written as:

@@

\CD{Just a proposal: put this subsection before Glisson Capsule}
We also combined our composite model with vascularization method proposed
by Peterl\'{i}k et al. \cite{Peterlik2012}. The method is based on linear beams
with local frames of reference (in a many aspects similar to the
corotational formulation). As such the model also handles geometric
non-linearities in the deformation. The proposed model assumes no relative
motion of vessels to the parenchyma. Segmented vascular network is
discretized into the set of linear elements that are then projected into
the tetrahedral mesh of the parenchyma. During the simulation forces and
torques resulting from the deformations of vessels by parenchyma are
computed and propagated back onto the tetrahedral mesh.

@@

\CD{If the experiments demonstrate that the stiffness of the material is different, the thickness is very very small... so we have to evaluate if it is really necessary to include the model of the capsule for both local and global deformations.}

@@

\CD{Maybe explain why we call that "local deformations" and in which case it is important to capture such local deformations (i.e. insertion of needle, contacts with a sharp instrument...).}

@@

\subsection{Global Deformations}
\CD{ Experiments demonstrate that capsule plays a role at local level, but still, maybe not so important to model the capsule for more global deformation...
+ What are the applications for modeling the global deformations of the liver (training and planning simulation, physics-based registration, augmented reality,)}
